---
title: "Introduction to RSQLite Database"
author: "Michelle Norris"
date: "November 11, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The website Kaggle.com has a collection of interesting datasets.  One is the World Development Indicators dataset at this url  [link](https://www.kaggle.com/worldbank/world-development-indicators).  Here is the description:

>"The World Development Indicators from the World Bank contain over a thousand annual indicators of economic development from hundreds of countries around the world. Here's a list of the available indicators along with a list of the available countries."

For example, this data includes the life expectancy at birth from many countries around the world.


Let's go to the url and check out a few cool graphs using this data.  

When you uncompress the data, you obtain many files.  The primary data are stored in a 1.4 GB file called **database.sqlite**.  There are auxiliary files which list information like: all countries in the database, all economic indicators in the database, notes about the data, etc. 

## Small Database Example 
Let's work with a small database first to get a feel for RSQLite.

```{r mtcars}
library(RSQLite)
myconnection = dbConnect(drv=SQLite(),":memory:") # sets up a temporary in-memory database

str(myconnection)
dbWriteTable(con=myconnection,name="mtcarsdata",value=mtcars)  # don't save this to a named object; mtcars is a built in dataset

dbListTables(myconnection)  #see all tables in myconnection
dbListFields(myconnection,"mtcarsdata")  #list fields=column names
```

## Basic SQL Query
A query is used to extract portions of a database or to summarize records in a database. The syntax of the query to extract specific columns is:  **SELECT column-list , FROM table-name** .  This is SQL (Structured Query Language) syntax and needs to be embedded in an R command for query.

```{r}
head(dbGetQuery(myconnection,"SELECT disp,cyl FROM mtcarsdata"))
head(dbGetQuery(myconnection,"SELECT * FROM mtcarsdata"))  #use the wildcard * to get all columns
```

## RSQLite with a Large database

Let's use RSQLite to read in the database, get information from the data, and run some queries.  First, put the database in your working directory.  Next, we must create a connection to the database.

```{r}
dbDisconnect(myconnection)  # disconnect the mtcars database
myconnection2 = dbConnect(drv=SQLite(), dbname="C:\\Users\\sac87931\\Documents\\kaggle\\world-development-indicators\\database.sqlite")  #connect R to the database
str(myconnection2)
dbListTables(myconnection2)
```


Each of the tables listed is a data frame.  Let's look at the fields (or column names in a few tables).

```{r}
head(dbListFields(myconnection2,"Indicators"))
head(dbListFields(myconnection2,"Country"))
```
## Extract Life Expectancy data by gender and Create a Graph

Now we extract data on life expectancy for females and males in the US and construct a time series graph.

```{r}

#female life expectancy at birth
results.female.life <- dbSendQuery(myconnection2, "SELECT * FROM Indicators WHERE CountryCode = 'USA' AND IndicatorCode='SP.DYN.LE00.FE.IN'" )

tempp = dbFetch(results.female.life, n=Inf)
head(tempp)
class(tempp)  # result is a dataframe

dbClearResult(results.female.life) # releases memory

plot(tempp$Year,tempp$Value,type="l",ylim=c(60,80),ylab="Life Expectancy",xlab='Year')


#male life expectancy at birth
results.male.life <- dbSendQuery(myconnection2, "SELECT * FROM Indicators WHERE CountryCode = 'USA' AND IndicatorCode='SP.DYN.LE00.MA.IN'" )
tempp = dbFetch(results.male.life, n=Inf)
head(tempp)
dbClearResult(results.male.life)
points(tempp$Year,tempp$Value,type="l",col="blue",lty=2)
```