---
title: "Term Project"
author: "Tanveer S Bariana"
date: "December 7, 2017"
output:
     html_document: 
       smart: false
---

In this R markdown file i will go through my results for the data science interveiw question for the prestigieous wikimedia foundation


```{r, echo= F}
wiki_data_raw <- read.csv(file = 'C:\\Users\\tanve\\Desktop\\dataSci\\stat196\\hmwrk\\termProject\\WikiData.csv')
library('data.table')
library('ggplot2')
wiki_data_raw <- data.table(wiki_data_raw)
```

here are the questions

1. What is our daily overall clickthrough rate? How does it vary between the groups?
2. Which results do people tend to try first? How does it change day-to-day?
3. What is our daily overall zero results rate? How does it vary between the groups?
4. Let session length be approximately the time between the first event and the last event in a session. Choose a variable from the dataset and describe its relationship to session length. Visualize the relationship.

so the folowing are my grpahs showing the overall daily click through rate and teh daily rates of Group A and B respectivly

```{r, echo= F}
wiki_data_dates <- wiki_data_raw[which(wiki_data_raw$action == 'visitPage' | wiki_data_raw$action == 'searchResultPage'),]

#cut out the section of the time stamp for hour, minute, or seconds so that you can get the dates. then make it a factor
wiki_data_dates$timestamp <- as.integer(wiki_data_dates$timestamp / 1000000)
wiki_data_dates$timestamp <- as.factor(wiki_data_dates$timestamp)
dates <- (unique(wiki_data_dates$timestamp))

#for each date fill a data table withs its number of clicks and searches based on groups
daily <- NULL
for (i in 1:length(dates)) {
    daily$Aclicks[i] <- length(which(wiki_data_dates$action == 'visitPage' & wiki_data_dates$timestamp == dates[i] & wiki_data_dates$group == 'a'))
    daily$Asearchs[i] <- length(which(wiki_data_dates$action == 'searchResultPage' & wiki_data_dates$timestamp == dates[i] & wiki_data_dates$group == 'a'))
    daily$Bclicks[i] <- length(which(wiki_data_dates$action == 'visitPage' & wiki_data_dates$timestamp == dates[i] & wiki_data_dates$group == 'b'))
    daily$Bsearchs[i] <- length(which(wiki_data_dates$action == 'searchResultPage' & wiki_data_dates$timestamp == dates[i] & wiki_data_dates$group == 'b'))
}
#use the data gathered above to get the averages below 
daily$Aavg <- daily$Aclicks / daily$Asearchs
daily$Bavg <- daily$Bclicks / daily$Bsearchs
daily$avg <- (daily$Aclicks + daily$Bclicks) / (daily$Asearchs + daily$Bsearchs)
daily <- data.frame(daily)
daily$date <- dates
#make the row names the dates for ease of understanding
rownames(daily) <- as.character(dates)
View(daily)
#graph teh findings 
clickthrough <- ggplot(data = daily, aes(x = daily$date, y = daily$avg))
clickthrough <- clickthrough + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.4)
clickthrough <- clickthrough + ggtitle('Over all daily Avg for each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
clickthrough
clickthroughA <- ggplot(data = daily, aes(x = daily$date, y = daily$Aavg))
clickthroughA <- clickthroughA + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.4)
clickthroughA <- clickthroughA + ggtitle('Daily Avg for group A on each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
clickthroughA
clickthroughB <- ggplot(data = daily, aes(x = daily$date, y = daily$Bavg))
clickthroughB <- clickthroughB + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.4)
clickthroughB <- clickthroughB + ggtitle('Daily Avg for group B on each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
clickthroughB

```

one can observe how the overal click rate and click rate for group A both decreases as time progresses but Group B increases  


now we move onto graphs for the Zero result rates, and the zero result rates of groups A and B respectivly 


```{r, echo= F}
#3 What is our daily overall zero results rate ? How does it vary between the groups ?
#new data table for new manipulation, similar process of setting it up as above
wiki_data_zero <- data.table(wiki_data_raw)
#make data set where its only searches with no results 
wiki_data_zero <- wiki_data_zero[which(wiki_data_zero$n_results == 0),]
wiki_data_zero$timestamp <- as.integer(wiki_data_zero$timestamp / 1000000)
wiki_data_zero$timestamp <- as.factor(wiki_data_zero$timestamp)
##for each date fill a data table withs its number of searches with no results
for (i in 1:length(dates)) {
    daily$Azero[i] <- length(which(wiki_data_zero$group == 'a' & wiki_data_zero$timestamp == dates[i]))
    daily$Bzero[i] <- length(which(wiki_data_zero$group == 'b' & wiki_data_zero$timestamp == dates[i]))
}
#use data gatered above to generate averages 
daily$Aavgzero <- (daily$Azero) / (daily$Asearchs)
daily$Bavgzero <- (daily$Bzero) / (daily$Bsearchs)
daily$avgzero <- (daily$Azero + daily$Bzero) / (daily$Asearchs + daily$Bsearchs)
View(daily)
#graph results
zeroRate <- ggplot(data = daily, aes(x = daily$date, y = daily$avgzero))
zeroRate <- clickthrough + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.8)
zeroRate <- clickthrough + ggtitle('Over all daily zero result rate for each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
zeroRate
zeroRateA <- ggplot(data = daily, aes(x = daily$date, y = daily$Aavgzero))
zeroRateA <- zeroRateA + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.8)
zeroRateA <- zeroRateA + ggtitle('Daily zero result rate for group A on each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
zeroRateA
zeroRateB <- ggplot(data = daily, aes(x = daily$date, y = daily$Bavgzero))
zeroRateB <- zeroRateB + geom_bar(aes(fill = daily$date), stat = 'identity', color = 'white', alpha = 0.8)
zeroRateB <- zeroRateB + ggtitle('Daily zero result rate for group B on each date') + theme_classic() + scale_fill_brewer(palette = "PuRd")
zeroRateB
```
we can observe that group A and B both have stable Zero Result rates very day with no major changes between days. WE can also note how the overall Zero Result rate decreases daily much liek how the overall daily click rate decreased